### 2023-12-06 scripts/manuscript.qmd
Final Touches & Abstract
Added the citations needed, updated figure outputs to be the same theme, added figure #s, and abstract.

---
title: "EDLD 651 Final Paper"
format: pdf
editor: visual
author: "Leah Peterson, Alexus Rock, and Ben Pfingston"
bibliography: biblio.bib
warning : false
message : false
echo : false
---

```{r}
#| include: FALSE
library(tidyverse)
library(here)
library(rio)
library(janitor)
library(ggplot2)
library(papaja)
library(gt)
library(gtExtras)
```

```{r}
#| echo: FALSE
communication <- import(here("data", "communication.csv"))
communication <- (clean_names(communication))
```

```{r}
#| echo: FALSE

#creating the new variables: the ppis subscale 1/parental invasive behaviors score, the ppis subscale 2/student defensive behaviors score, and the suspicion scale score
comm <- communication %>%
  mutate(ppis_1_score = rowSums(communication[,5:7])) %>%
  mutate(ppis_2_score = rowSums(communication[,8:15])) %>%
  mutate(ss_score = rowSums(communication[,16:27]))

comm <- comm %>%
  mutate(ppis_total = ppis_1_score + ppis_2_score,
         age = as.numeric(age)) %>%  # Converting age column to numeric
  filter(!is.na(age))
```

```{r}
#| echo: FALSE

#pivoting the data, also removing unnecessary columns
communication <- pivot_longer(comm,
                      cols = ppis_1_1:ss_12,
                      names_to = "item",
                      values_to = "score")

communication <- select(communication, -c(2:7))
```

```{r}
ppis_ss_table <- communication %>%
select(ppis_1_score:ppis_total) %>%
  pivot_longer(
  cols = everything(),
  names_to = "measure",
  values_to = "score") %>%
  group_by(measure) %>%
  summarise(Mean = round(mean(score, na.rm = TRUE), 2),
            SD = round(sd(score, na.rm = TRUE),2)) %>%
  pivot_longer(
    cols = -measure,
    names_to = "desc",
    values_to = "value"
    ) %>%
  pivot_wider(names_from = measure,
              values_from = value)
```

# Abstract

By drawing on theories of gender socialization and trust violations, the current study explores the dynamics of trust and privacy invasions in student-teacher relationships. Additionally, this study aims to investigate how gender may impact a young adult's perception of these topics. We predict that individuals who are more trustworthy will perceive fewer privacy invasions from teachers in their life. A sample of 98 young adults from a Minnesota undergraduate student body participated in this study. To measure young adult perceptions of trustworthiness, prior teacher privacy invasions, and suspicion towards teachers, modified scales, including the Parental Privacy Invasions Scale and the Suspicion Scale, were used. The results indicate that self-reported trustworthiness does not significantly impact the perception of teacher privacy invasions. However, gender differences were observed in how individuals perceive and experience their teacher behaviors. This study adds to our understanding of the complex interplay between trust, gender, and privacy invasions in student-teacher relationships. Further, the findings emphasize the complexity of trust dynamics in student-teacher relationships and suggest young adults perceive and respond to invasive behaviors in educational settings differently based on their gender identity.

\newpage

# Introduction

By the time a child turns 18, they have spent approximately 15% of their life at school [@wherry2004]. Thus, for many young adults, teachers they've encountered are often some of the most influential people in shaping their current knowledge base. In addition to influencing student learning and test-scores, supportive teacher communication behaviors have been shown to impact feelings of satisfaction and self-efficacy [@blazar2016; @frymier2000; @chesebro2000]. Additionally, @Bruney2013 found that teacher relationships are crucial for a young adult's independent thinking skills, which can lead to a greater life satisfaction. While childhood is often associated with growth and development, young adulthood is also a time of great advancement in life [@parks2005]. During this period, young adults are beginning to apply the knowledge they have gained throughout their lives without direct guidance from a parent or authority figure. As such, to be successful in this stage of life, young adults need a strong foundation in independent thinking [@Bruney2013]. In order to build that foundation, experiences of trustful interpersonal relationships with authority figures, such as teachers, are essential in providing support as young adults reconfigure their identities [@parks2005]. Yet, there is a dearth of research focused on investigating the ways in which overall trust of young adults may be influenced by teachers' communicative behaviors.

## (Dis)Trust & Invasive Behaviors

Trust is a multifaceted concept that encompasses elements such as honesty, connectedness, reliability, and competency [@simpson2007; @raider-roth2005]. Further, trust is not an automatic feature in a relationship. Caring for and emotional investment in interactions has been shown to be an important factor for building trust [@kuhlmann2008; @smith2005]. When there is trust in a relationship, commonly individuals act consistently, follow through, and work to meet expectations [@lewicki2006]. Alternatively, when trust has not been cultivated in a relationship, individuals exhibit greater anti-social behaviors and have impaired psychological adjustment [@simpson2007]. Additionally, trust in relationships can be broken. Specifically, broken trust that is accompanied by invasive behaviors or deception creates the most harm in a relationship and that trust may never fully recover [@schweitzer2006a].

According to @ledbetter2012, invasive behaviors are those that intend to violate an individual's privacy. There are three main categories for invasive behaviors: subversive or spatial invasions, direct or verbal invasions and mediated invasions [@ledbetter2012; @petronio2002]. Subversive or spatial invasions are acts such as eavesdropping and listening to phone conversations or searching through belongings [@ledbetter2012; @petronio2002]. Direct or verbal invasions are when something is said directly to the individual who is keeping their information private [@ledbetter2012; @petronio2002]. Lastly, mediated invasions are when an investigation for information happens through going on someone's social media account, reading emails or text messages or tracking phone activity [@ledbetter2012]. When trust is broken, individuals commonly are less likely to maintain or create relationships with others in the future, experience anger, become less cooperative in group settings, and give negative evaluations of others [@ledbetter2012; @lount2008; @simpson2007]. Thus, it is hypothesized that individuals who are more trustworthy will perceive less privacy invasions.

## Trust & Gender

Socialization perspectives on gender differences propose that women are more inclined to define themselves in relational terms compared to men [@cross1997]. This inclination makes women particularly motivated to cultivate positive and nurturing relationships [@cross2000; @amanatullah2008]. In the context of social relationships, trust has commonly been viewed as the "glue" that holds the relationship together [@rotenberg1995]. Additionally, following a trust transgression, women are more likely than men to maintain trust in the relationship [@haselhuhn]. Thus, it is hypothesized that there will be differences in how female, male, and non-binary individuals perceive their own levels of trust towards others, and how they experience privacy invasions.

# Method

## Sample

```{r}
mean_age <- mean(comm$age) %>% round(2)
sd_age <- sd(comm$age) %>% round(2)
```

98 participants were recruited primarily through an undergraduate student body located in Minnesota. The participation criteria required that participants be at least 18 years old (*M =* `r mean_age`, *SD =* `r sd_age`) and have interacted with a teacher in a school setting. Demographic information, such as gender identity and ethnic-racial identity, was also collected.

## Procedure

Participants were provided with a link to the survey, which was conducted through the online system Qualtrics. Before the survey began, each participant was given an online consent form that provided information about the length of the survey, confidentiality, their ability to withdraw from the study at any time and contact information for the researchers.

## Measures

To measure different aspects of student-teacher trust interactions, items were modified from the @ledbetter2012 Parental Privacy Invasions Scale (PPIS) and the @levine1991 Suspicion Scale.

### Modified Parental Privacy Invasions Scale: Teacher Privacy Invasions Scale (TPIS)

Eleven items were adapted from the Parental Privacy Invasions Scale to reflect teacher, rather than parental, interactions. Each item was measured on a five-point Likert scale, ranging from 0 to 4, with response options including, "Never", "Rarely", "Often", "Sometimes", and "Always".

Three items were modified the Parental Invasive Behavior subscale of the PPIS to measure student perceptions of invasive behaviors from their teachers. One example item from this subscale is, "My teachers demand that I change my behavior in some area of my life."

The other 8 modified items came from the Children's Defensive Behavior subscale, which captured the young adults' self-reports of behaviors to safeguard their privacy from their teachers. An example of an item from this subscale is, " I avoid going to see my teacher outside of normal school/class hours. (such as before or after class, during break, before or after school tutoring)."

### Modified Suspicion Scale (SS)

Twelve items were adapted from the Suspicion Scale to capture young adults' suspicion toward their teachers. Each item was measured on a seven-point Likert scale, ranging from 0-6, with response options including, "Strongly Disagree", "Somewhat Disagree", "Disagree", "Neither Agree nor Disagree", "Somewhat Agree", "Agree", and "Strongly Agree". Example items from the scale include, "Basically I am a trusting person", and "Most teachers follow the saying"honesty is the best policy."

# Results

```{r}
model <- lm(ppis_total ~ ss_score, comm)
modelresults <- apa_print(model)
slope <- coef(model)[2] %>% round(2)
```

We used R (Version 4.3.2; @R) and the R-packages *ggplot2* [@ggplot], *gt* [@gt], *gtExtras* [@gtE], *here* [@here], *janitor* [@janitor], *papaja* [@papaja], *rio* [@rio], and *tidyverse* [@tidy] for all our analyses.

## Descriptive Statistics

The means and standard deviations of the TPIS scores and the SS Scores can be shown in Table 1 below.

\newpage

```{r}
#| output: asis
desc_table <- 
  gt(ppis_ss_table) %>%
  cols_label(desc = "Statistic", ppis_1_score = "Teacher Privacy Invasions", ppis_2_score = "Student Trustworthiness", ppis_total = "TPIS Total", ss_score = "Suspicion Scale") %>%
  tab_caption(caption = "Descriptive Statistics of TPIS and SS Scores ") %>%
  tab_footnote(footnote = "Table 1.")

desc_table
```

## Teacher Privacy Invasions and Suspicion Scale

Privacy invasions and self-rated trustworthiness were not found to be correlated. Self-rated trustworthiness was not a significant predictor of privacy invasions (`r modelresults$full_result$ss_score`). A scatterplot of the relationship can be seen in Figure 1, shown below.

```{r}
#| label: scatterplot

ggplot(comm, aes(ss_score, ppis_total)) +
  geom_point() +
  geom_smooth(method = "lm", color="darkgreen") +
  labs(title = "Student's Trustworthiness and Teacher Privacy Invasions", 
       x = "Own Precieved Trustworthiness",
       y = "Teacher Privacy Invasions",
       tag = "Figure 1.") +
   geom_text(x = 30, y = 4,
     label = sprintf("Slope: %.2f", coef(lm(ppis_total ~ ss_score, data = comm))[2])) +
  theme_minimal()
```

## Gender Differences

Gender differences existed for teacher invasive behaviors. Non-binary individuals perceived their teachers behaviors as being less invasive than females and males perceptions. Gender differences existed for teacher privacy invasion. Males perceived their teachers invading their privacy less than female and non-binary individuals. A boxplot of Teacher Invasive Behaviors scores, split by gender, can be found in Figure 2. There were no gender differences in student defensive behaviors. A boxplot of this relationship can be found below in Figure 3. There were also no differences in the total Teacher Privacy Invasion scale scores between gender groups, shown in the boxplot in Figure 4.

There were gender differences in the amount of suspicion towards teachers. Females had the highest levels of suspicion towards their teachers, and non-binary individuals had the least amount of suspicion. A boxplot of mean Suspicion Scale scores, split by gender, can be found in Figure 5 below.

```{r}
#| label: boxplot-gender-ppis-1-score

ggplot(communication, aes(gender, ppis_1_score, fill=gender)) +
  geom_boxplot() +
  theme(legend.position = "none") +
    scale_x_discrete(limits = c("Female", "Male", "Non-binary")) +
  labs(x = "Gender",
       y = "Teacher Invasive Behaviors Score",
       title = "Boxplot of Teacher Invasive Behaviors Score by Gender", 
       tag = "Figure 2.")+
  theme_minimal()
```

```{r}
#| label: boxplot-gender-ppis-2-score
ggplot(communication, aes(gender, ppis_2_score, fill=gender)) +
  geom_boxplot() +
  theme(legend.position = "none") +
  scale_x_discrete(limits = c("Female", "Male", "Non-binary")) +
  labs(x = "Gender",
       y = "Student Defensive Behaviors Score",
       title = "Boxplot of Student Defensive Behaviors Score by Gender",
       tag = "Figure 3.") +
  theme_minimal()

```

```{r}
#| label: boxplot-gender-ppis-total

ggplot(communication, aes(gender, ppis_total, fill=gender)) +
  geom_boxplot() +
  theme(legend.position = "none") +
    scale_x_discrete(limits = c("Female", "Male", "Non-binary")) +
  labs(x = "Gender",
       y = "Teacher Privacy Invasion Scale Total",
       title = "Boxplot of Teacher Privacy Invasion Scale Total by Gender",
       tag = "Figure 4.") +
  theme_minimal()

```

```{r}
#| label: boxplot-gender-ss-score
ggplot(communication, aes(gender, ss_score, fill=gender)) +
  geom_boxplot() + 
  scale_x_discrete(limits = c("Female", "Male", "Non-binary")) +
  theme(legend.position="none") +
  labs(x = "Gender",
       y = "Suspicion Scale Score",
       title = "Boxplot of Suspicion Scale Score by Gender",
       tag = "Figure 5.") +
  theme_minimal()

```

# Discussion

In this study, self-reported trustworthiness had no influence on perceived privacy invasions. Individuals who were more trustworthy did not perceive less privacy invasions than those with less trustworthiness. Levels of trustworthiness has no influence on how students perceive their privacy being invaded by teachers.

There were gender differences in how female, male, and non-binary individuals perceived their own levels of trust towards others, and how they experienced privacy invasions. Non-binary individuals perceived their teacher's behaviors as being less invasive than female and male individuals. Males perceived their teachers invading their privacy less frequently than female and non-binary individuals. Females had the highest levels of suspicion towards their teachers, and non-binary individuals had the least amount of suspicion.

There were no gender differences in student defensive behaviors. This indicates that students were not more defensive towards their teachers depending on student gender.

\newpage

# References